---
title:  'Week 7 - R Vector/matrix manipulations and math, speed considerations'
author: "Robert Settlage"
date: "2021-10-12"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
header-includes: \setlength\parindent{24pt} \usepackage{MnSymbol} \usepackage{mathrsfs}
---


```{r setup, eval=T, echo=F, include=F}
  library(ggplot2); library(microbenchmark)
```

## Today's Agenda

- Review good programming practices  
- R functions  
- matrix operations 

## Good Programming Practices

Google's R Style Guide <https://google.github.io/styleguide/Rguide.xml>  
Hadley Wickam's Style Guide: <http://r-pkgs.had.co.nz/style.html>  

Remember: ultimate goal is Reproducible Research.  
To enable this, we need well documented (annotated) _**readable**_ code.

1. variable/object names  
    + nouns describing what the object holds, e.g. originaData instead of d  
    + DO NOT use existing variables or functions, e.g. T<-1000
2. function formation  
    + function names should be verb associated with the function purpose  
    + comment both function purpose and required arguments    
    + arguments should have defaults (my personal preference)
    + my preference is to explicitly return a value  
3. commenting rules  
    + comment your code!!  'nough said?  
4. indent within logical blocks of code !!
    + indenting improves the readability of the code by orders of magnitude!  
    
## Functions

A function is an _**object**_ that takes some input objects (args) and produces some output objects.  _**All**_ work in R is done in functions. '[<-'(animals,4,"duck")

```{r echo=T, eval=F}
    dataA<-1;dataB<-2

    BAD<-function(x){
    mean(c(x,dataA))
    }
    
    computeMeanGOOD<-function(x=3,y=5){
        # quick function to compute the geometric mean of two numbers, x and y
        # returns geometric mean
        if (!is.numeric(x) || !is.numeric(y)) stop("both x and y must be numeric")
        return((x+y)/2)
    }
    
    BAD(x=dataB)
    computeMeanGOOD(x=dataA,y=dataB)
```

GOOD functions include passing in ALL necessary data.  If we do not do this, we can have MAJOR issues with variable scope that is often difficult to troubleshoot.  Ideally our code also includes error checking.

## Variable scope

R uses Lexical scoping rules.  What tha?

<https://darrenjw.wordpress.com/2011/11/23/lexical-scope-and-function-closures-in-r/>  
Totally recommend subscribing to Darren's blog AND reading all his past posts.  

<http://andrewgelman.com/2014/01/29/stupid-r-tricks-random-scope/>  
Another good site.  Also has a few great stats books.

## Timing

As you start to make functions, it is a good idea to think about how long things take.

Really good blog on this here:  
<https://www.r-bloggers.com/5-ways-to-measure-running-time-of-r-code/>


## Bit more on functions in R 

<https://csgillespie.github.io/efficientR/programming.html>

Everything in R is essentially a function call.  Generally, this means at some point the function call will make a call to the underlying C/Fortran code.

"A golden rule in R programming is to access the underlying C/Fortran routines as quickly as possible; the fewer functions calls required to achieve this, the better. For example, suppose `x` is a standard vector of length `n`. Then:"

## Memory allocation -- Method 1

Memory management is the Achilles' heel of many programming languages.  If we are careful in our programming, we can alleviate many of the issues.

Consider the following examples given in Gillespie's R Book:

Method 1: create empty vector, grow vector to final size

```{r echo=T, eval=T, include=T}

method1 = function(n) {
  vec = NULL # Or vec = c()
  for(i in seq_len(n))
    vec = c(vec, i)
  vec
}
```

## Memory allocation -- Method 2

Method 2: create vector of final length and change values via subscripting

```{r echo=T, eval=T, include=T}
method2 = function(n) {
  vec = numeric(n)
  for(i in seq_len(n))
    vec[i] = i
  vec
}
```

## Memory allocation -- Method 3

Method 3: create final object directly

```{r echo=T, eval=T, include=T}
method3 = function(n) seq_len(n)
```

Thoughts on the 3 methods??

```{r echo=T, eval=F, include=T}
library(ggplot2)
library(microbenchmark)
times<-microbenchmark(temp1=method1(10000),temp2=method2(10000),temp3=method3(10000),times = 100, unit = "ms",control=list(warmup=0))
ggplot2::autoplot(times)
```


## Mean is sloooooow

```{r echo=T, eval=F, include=T}
N <- 100000 #### <- might start a little lower
#manual vs internal mean function
microbenchmark(result1<-sum(1:N)/N,
               result2<-mean(1:N),times = 100, unit = "ms")
```


## Memory and functions

This topic is really intertwined with lexical scoping.  The basic idea is that R manages objects by creating environments.  Each environment is essentially a bag of names pointing to thier associated memory space.  Creating a function, creates a new environment.  The environment lives as long as the function is active.

<http://adv-r.had.co.nz/Functions.html#lexical-scoping>  
<http://adv-r.had.co.nz/Environments.html>

As you program in R more, you will want to pay special attention to memory management and function calls.

-- memory used  
-- speed issues  

look at memory size of objects
look at memory use of R
look at memory use of R during a function call with large object

## R and useful matrix stuff

1. matrix mult %*%
2. transpose t()
3. create diagonal matrix diag()
    + this one has two cases
4. matrix inverse solve()
5. determinate det()
6. rank matA <- qr(A); matA$rank
7. t(x) %*% y crossprod()

More comprehensive list here:
<https://www.statmethods.net/advstats/matrix.html>


## vectors and matrices in R

Super easy to create:

```{r echo=T, eval=F, include=T}
    a <- c(1:5)
    B <- matrix(1:10,ncol=2)
    a
    B
```

## basic operations

```{r echo=T, eval=F, include=T}

## add scalar
a
a + 5
## multiply by scalar
B
B * 5
## multiply vectors/matrix
a * a
a * B
B * a
## length of vector
sqrt(sum(a*a))
```

## matrix math

```{r echo=T, eval=F, include=T}
## transpose
t(a)
t(t(B))
## matrix multiply
a %*% B
B %*% a
t(B) %*% a
## diag
diag(1,3)
diag(c(1,2,3))
diag(B)
## solve
C <- matrix(1:9 + rnorm(9,0,1e-3),ncol=3) 
C %*% solve(C)
## trace
sum(diag(C))
## det
det(C)
## Kronecker product
I <- diag(3)
J <- matrix(1,3,3)
blockDiag <- kronecker(I, J)
```

## Using duality to do stuff

colMeans.  What if we want the means of a matrix by columns.  Of course there are functions for this, but, for kicks, can we do this with some mixed matrix art ... ??

```{r eval=F, echo=T, include=T}

C <- matrix(1:9 + rnorm(900,0,1e-3),ncol=30)
colMeans(C) ## presumably optimized version
ones <- rep(1,30)
t(ones) %*% C / 30 ## manual but using matrix operations

library(microbenchmark)

#for kicks
microbenchmark(result1<-{colMeans(C)},
               result2<-{t(ones) %*% C / 30},times = 100, unit = "ms")

```

## Making things faster:

A lot of effort has gone into making matrix math faster on computers.  Again, this is a guided tour, so I will show you a few things and hope this sparks your curiousity enough to hit google:

```{r echo=T, eval=F, include=T}

    A = matrix(rnorm(20*40000, mean=0, sd=5), 20, 40000)
    B = matrix(rnorm(20*40000, mean=0, sd=5), 20, 40000)
    time1 <- system.time({t(A)%*%B})
    time2 <- system.time({crossprod(A,B)})
    print(rbind(time1,time2))

```

Don't invert matrices:

<https://www.johndcook.com/blog/2010/01/19/dont-invert-that-matrix/>
<https://www.r-bloggers.com/dont-invert-that-matrix-why-and-how/>

```{r echo=T, eval=F, include=T}

    n <- 5000
    A = matrix(rnorm(n*n, mean=0, sd=5), n, n)
    x <- rnorm(n)
    system.time({b_inverting <- solve(A)%*%x})
    system.time({b_solving <- solve(A,x)})

    max(abs(b_inverting - b_solving))
```

## For those interested in boosting their local R

Not all installations of R come accelerated:  

*There are several highly optimized libraries that can be used instead of the default base libraries. These libraries are optimized to take advantage of the hardware they are run on, and can be significatantly faster than the base implementation (operations such as Matrix multiplications may be over 40 times faster*  
<https://csantill.github.io/RPerformanceWBLAS/>

