---
title:  'Formula interface and multi panel plots in ggplot2'
author: "Robert Settlage"
date: "2021-09-20"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
header-includes: \setlength\parindent{24pt} \usepackage{MnSymbol} \usepackage{mathrsfs}
---

## Today's Agenda

- Review: Multipanel plots and ggplot2  
- Formula interface  
- Multipanel ggplot2 plots

### Data wranging -- common plots

1. histogram
2. box plot vs violin plot
3. scatter plot
4. pie chart, ...

*don't forget color and shape as dimensions*

## Plot inspiration

<http://www.r-graph-gallery.com>


### Multipanel base plot

Multipanel plots in base R means manipulating the canvas.

This site has everything needed to make fantastic plots using base functions:

<https://www.statmethods.net/advgraphs/layout.html>

### Multipanel base plot example

<https://r-charts.com/base-r/combining-plots/>

```{r multipanel_example, echo=T, eval=T, include=T}
# Data
set.seed(6)
x <- rexp(50)

layout(matrix(c(2, 0, 1, 3), nrow = 2, ncol = 2, byrow = TRUE),
       widths = c(3, 1),
       heights  = c(1, 3), respect = TRUE)

# Top and right margin of the main plot
par(mar = c(5.1, 4.1, 0, 0))
plot(x, cex=2, pch=20)

# Left margin of the histogram
par(mar = c(0, 4.1, 0, 0))
hist(x, main = "", bty = "n", axes = FALSE, ylab = "")

# Bottom margin of the boxplot
par(mar = c(5.1, 0, 0, 0))

# Boxplot without plot region box
par(bty = "n")

# Boxplot without axes
boxplot(x, axes = FALSE)
```

## Margins are the key

https://www.r-graph-gallery.com/74-margin-and-oma-cheatsheet.html

```{r margins, echo=TRUE, include=TRUE, eval=TRUE}
# Margins area
par(oma=c(3,3,3,3)) # all sides have 3 lines of space
par(mar=c(5,4,4,2) + 0.1)

# Plot
plot(0:10, 0:10, type="n", xlab="X", ylab="Y") # type="n" hides the points

# Place text in the plot and color everything plot-related red
text(5,5, "Plot", col="red", cex=2)
box(col="red")

# Place text in the margins and label the margins, all in forestgreen  
mtext("Margins", side=3, line=2, cex=2, col="forestgreen")  
mtext("par(mar=c(b,l,t,r))", side=3, line=1, cex=1, col="forestgreen")  
mtext("Line 0", side=3, line=0, adj=1.0, cex=1, col="forestgreen")  
mtext("Line 1", side=3, line=1, adj=1.0, cex=1, col="forestgreen")  
mtext("Line 2", side=3, line=2, adj=1.0, cex=1, col="forestgreen")  
mtext("Line 3", side=3, line=3, adj=1.0, cex=1, col="forestgreen")  
box("figure", col="forestgreen")  
 
# Label the outer margin area and color it blue  
# Note the 'outer=TRUE' command moves us from the figure margins to the outer margins.  
mtext("Outer Margin Area", side=1, line=1, cex=2, col="blue", outer=TRUE)  
mtext("par(oma=c(b,l,t,r))", side=1, line=2, cex=1, col="blue", outer=TRUE)  
mtext("Line 0", side=1, line=0, adj=0.0, cex=1, col="blue", outer=TRUE)  
mtext("Line 1", side=1, line=1, adj=0.0, cex=1, col="blue", outer=TRUE)  
mtext("Line 2", side=1, line=2, adj=0.0, cex=1, col="blue", outer=TRUE)  
box("outer", col="blue") 
```

### Elements of a good figure | EDA stage

1. simple to create  
2. shows only what you need it to
3. doesn't take a lot of explaination  
4. YOU are the primary audience (at first)  
5. has enough detail to give insight

### That always means exploring the available factors and relationships

```{r boxexample, echo=F, include=T, eval=T, fig.width=6, fig.height=4}
    library(beeswarm)
    sex <- (runif(1000)>0.5)*1
    drug_duration <- sex*0.1 + rnorm(1000, 0, 0.1) + 5
    drug_info <- data.frame(cbind(drug_duration,sex))
    drug_info$id <- ifelse(drug_info$sex==0,"M","F")
    par(mfcol=c(1,2))
    boxplot(drug_info$drug_duration)
    beeswarm(drug_duration,add=T, pch=20, cex=0.3)
    boxplot(drug_duration~id,drug_info)
    beeswarm(drug_duration~id,drug_info, add=T, pch=20, cex=0.3)
```

### Elements of a good figure | Publication stage

1. ok, wow factor  
2. shows only what you need it to  
3. draws the reader in  
4. invokes inquiry  
5. is fully self contained with legend AND caption 
6. simple is GENERALLY better (meaning less factors in play)  
7. all these rules are out the window for art


### ggplot2

Another Hadley Wickam creation.  Lot's of love/hate for it.  I mostly like it.  I find a few types of plots are SUPER simple via ggplot2 that might be very time consuming in base graphics.  I also find that I spend more time in customizing away the ggplot look if I am trying to make a figure for a pub.

<https://www.sharpsightlabs.com/blog/geom_line/?utm_source=ssl_email_primary&utm_medium=email&utm_campaign=newsletter>

The basic idea is that creating figures is a step wise process of building a plot.  

1. you need data  
2. you need a mapping of data to the plot features  
3. you need a graphical object  

You mix these and wallah!!

And, like everything else, people are opinionated ...

<https://flowingdata.com/2016/03/22/comparing-ggplot2-and-r-base-graphics/>  
<https://simplystatistics.org/2016/02/11/why-i-dont-use-ggplot2/>  
<http://varianceexplained.org/r/why-I-use-ggplot2/>  

### Example

Quick example of a basic plot using both base graphics and ggplot2

```{r base_r, echo=TRUE, eval=TRUE, include=TRUE}

par(oma=c(0.5,1,0.5,1))
plot(x=mtcars$mpg, y=mtcars$disp, pch=20, col=as.factor(mtcars$gear),
     xlab="mpg", ylab="disp", main="Fuel economy declines as displacement increases")
legend("topright", legend=levels(as.factor(mtcars$gear)),
       title="gears",col=1:3,pch=20)

```

```{r ggplot_mtcars, echo=TRUE, eval=TRUE, include=TRUE}
library(ggplot2)
ggplot(data=mtcars, aes(x=mpg, y=disp, color=as.factor(gear))) + 
  geom_point() + 
  theme(panel.background = element_rect(fill = "white", colour = "grey50"),
        legend.position = c(0.85, 0.85)) + 
  labs(color = "gears", 
        title = "Fuel economy declines as displacement increases")
```

## Formula interface

Consider the following data:

```{r echo=F, eval=T, include=T}
    sex <- (runif(1000)>0.5)*1
    drug_duration <- sex*0.1 + rnorm(1000, 0, 0.1) + 5
    drug_info <- data.frame(cbind(duration=drug_duration,sex))
    drug_info$age <- round(runif(1000,min=20,max=60)+drug_info$sex*drug_duration*0.2,0)
    drug_info$sex <- ifelse(drug_info$sex==0,"M","F")
    drug_info$hair_color <- sample(c("blond","brown","red"), size=1000, 
                                   replace=TRUE, prob=c(0.4,0.4,0.2))
    knitr::kable(head(drug_info),caption="drug action")
```

Given this data, we might envision wanting to evaluate relationships amonst the variables to answer questions like how does duration of drug respons change based on sex, or age, or sex and age.  How do we capture that in a simple way???

### Formulas

Formulas in R are objects.  We create them like any other object, through assignment.

```{r formula_examples, eval=TRUE, include=TRUE, echo=TRUE}
f1 <- drug_time ~ sex
f2 <- drug_time ~ sex + age
class(f1)
```

The variables on both sides of the `~` are named like in your linear models class -- `dependent variable ~ independent variable`

### Sided formulas 

Note that you can have 1- and 2-sided formula:

+ 1-sided  
  always right side variables, ie independent
  `~ x`

+ 2-sided
  `y ~ x`
  
### Formula operators

+ `+` read as and  
+ `-` read as minus
+ `|` read as conditioned on  
+ `:` interaction between  
+ `*` crossing  of variables, including the interaction  
+ `%in%` nesting  
+ `^` to limit crossing  
+ `I()` as-is, useful for example when including `x^2`
+ `.` means all

A few functions of formula: terms, all.vars, update

```{r formula_functions, eval=TRUE, echo=TRUE, include=TRUE}
myformula <- y ~ (a + b + c)^2
terms(myformula)
update(myformula,"~.+d*e")
```

## Formula interface and figures

Many plotting functions take formula.  The ggformula package extends ggplot2 to also take formula.

```{r plot_formula, echo=TRUE, eval=TRUE, include=TRUE}

plot(duration~age, data=drug_info, col=as.factor(sex), pch=20)
boxplot(duration~sex, data=drug_info)
library(beeswarm)
beeswarm(duration~sex, data=drug_info, add=T, cex=0.3, pch=20, col=sex+1)
library(ggformula)
gf_point(duration~age, data = drug_info, color=~sex, size=~log(duration)-1)
```

### Note on code chunk output

Do you REALLY want all the ggformula load info in your report?  Consider having a setup chunk.

Also note the options that are set.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, include=TRUE, eval=TRUE)
library(ggformula)
```

```{r test}
gf_point(duration~age, data = drug_info, color=~sex, size=~log(duration)-1)
```


## GGplot2 and multipanel

Multipanel plots using ggplot.  

First, if what you want is essentially similar plot, but seperated by a factor level, consider faceting.

Second, if instead, you want to merge dissimilar plots into a single figure, consider ggarrange (ggpubr) or a similar canvas controlling package (gridExtra, patchwork, ...).

There is, of course, a book... <https://ggplot2-book.org/arranging-plots.html>

```{r ggplot_multipanel}
ggplot(data=drug_info, aes(x=age, y=duration)) + 
  geom_point() +
  facet_grid(sex~hair_color)
```

I will leave the other packages to you.

## Next up

+ Homework 3
