---
title: "Homework 2"
subtitle: "Due: Next Wednesday, September 15, 2021"
author: "Robert Settlage"
date: "2021-09-08"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
header-includes: \setlength\parindent{24pt} \usepackage{MnSymbol} \usepackage{mathrsfs}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

For each assignment, turn in by the due date/time.  Late assignments must be  arranged prior to submission.  In every case, assignments are to be typed neatly using proper English in Markdown.

This week, we spoke about Reproducible Research, R, Rstudio, Rmarkdown, and LaTeX.  To summarize the ideas behind Reproducible Research, we are focusing on Reproducible Analysis.  For us, Reproducible Analysis is accomplished by mixing code, figures and text into a cohesive document that fully describes both the process we took to go from data to results and the rational behind our data driven conclusions.  Our goal should be to enable a moderately informed reader to follow our document and reproduce the steps we took to reach the results and hopefully conclusions we obtained.

# Problem 1 -- Rstudio.cloud Primers

R is an open source, community built, programming platform.  Not only is there a plethra of useful web based resources, there also exist in-R tutorials.  Please check out both of the Primers labeled as The Basics on Rstudio.cloud.

For those of you exploring Julia, please check out the basics and plotting tutorials here: 
<http://ucidatascienceinitiative.github.io/IntroToJulia/Html/BasicIntroduction>  
<http://docs.juliaplots.org/latest/>

# Problem 2

Now that we have the R environment setup and have a basic understanding of R, let's add Markdown (choose File, New File, R Markdown, pdf).  

Let's go ahead and save the file as is.  Save the file to the directory containing the *README.md* file you created and committed to your git repo in Homework 1.  The filename should be: HW1_pid, i.e. for me it would be HW2_rsettlag.  Note, in future homework assignments I -may-, ok, *will likely* forget to increment the number, please infer the proper assignment number from the assignment...

You will use this new R Markdown file for the remainder of this homework.

To start, let's prove we can type some basic $LaTeX$ text and function elements:

### Part A
In this new Rmarkdown file, please type a paragraph about what you are hoping to get out of this class.  Include at least 3 specific desired learning objectives in _list_ format.  

### Part B
To this, add 3 density functions (Appendix Cassella & Berger) in centered format with equation number, i.e. format this as you would find in a journal.  


# Problem 3

A quote from Donoho (1995): "an article about computational results is advertising, not scholarship. The actual scholarship is the full software environment, code and data, that produced the result."  To the document created in Problem 2, add a summary of the steps in performing Reproducible Research in numbered list format as detailed in:  
<http://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1003285>.  
\newline
Next to each item, comment on any challenges you see in performing the step.  If you are interested in learning more, a good summary of why this is important can be found in  
- <https://www.informs.org/ORMS-Today/Public-Articles/October-Volume-38-Number-5/Reproducible-Operations-Research>  
- <https://doi.org/10.1093/biostatistics/kxq028>  
- <http://statweb.stanford.edu/~wavelab/Wavelab_850/wavelab.pdf>

# Problem 4

Here we will create a linear model and associated basic diagnostic plots for some COVID data.  What we want is US data only and we will look at the 14 day moving average per 100k people.  We are going to limit the time period to June-July 2020.

```{r libraryLoad, echo=TRUE, include=TRUE, eval=FALSE}
install.packages('data.table')
library(data.table)
covid_raw <- fread("https://opendata.ecdc.europa.eu/covid19/casedistribution/csv")
us <- covid_raw[covid_raw$countriesAndTerritories == 'United_States_of_America',]
us_filtered <- us[us$month %in% c(6:7),]
us_filtered$index <- rev(1:dim(us_filtered)[1]) 
fit<-lm(`Cumulative_number_for_14_days_of_COVID-19_cases_per_100000`~index, data=us_filtered)
```

### Part A
1. Create a summary table of the us_filtered data.  Hint, use `summary` and `kable` in the knitr package.  How many time points have we limited ourselves to?  Are there missing values?

2. Create a summary table of the linear model output.  Hint, use the stargazer package.

### Part B
Here we need to create a few diagnostic plots.  We could use the plot function on the fitted model, ie `plot(fit)` or `plot(fit,1)`, but let's create them manually.  Start by augmenting the model with diagnostics:

```{r augment_model, eval=FALSE, echo=TRUE, include=TRUE}
install.packages("broom")
fit.diags <- broom::augment(fit)
```

We are going to restrict this to four plots:  

1. residuals vs fitted  
2. normal Q-Q  
3. scale-location  
4. residuals vs leverage  

To see how these should look, look at plot numbers 1-3, and 5 using plot(fit,number).

### Part C
Create an auto correlation plot of the residuals using the acf function.  The pattern observed is indicative of a time pattern.  The simple linear model is not appropriate in this case.

# Problem 5

Combine the plots from Problem 4B into a single plot.  Pay attention to the margins and reduce the margins such that the figure plot area is maximized compared to margins.


**This document containing solutions to Problems 2-4 should be typed in RMarkdown, using proper English, and knitted to create a pdf document.  Do NOT print, we will use git to submit this assignment as detailed below.**

# Problem 6
Please knit this document to PDF (name should be `HW2_pid`) and push to GitHub:

In the R Terminal, type:  

* git pull  
* git add HW2\_pid.[pR]*  (NOTE: this should add two files)  
* git commit -m "final HW2 submission"   
* git push

A more detailed description is on the course website under *Submitting Homework*.

## Reminder on where to find Git help:  
Read through the Git help Chapters 1 and 2. <https://git-scm.com/book/en/v2>






