---
title: "Homework 3"
subtitle: "Due: Wednesday, September 29, 2021"
author: "Robert Settlage"
date: "2021-09-08"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
header-includes: \setlength\parindent{24pt} \usepackage{MnSymbol} \usepackage{mathrsfs}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggformula)
```

For each assignment, turn in by the due date/time.  Late assignments must be  arranged prior to submission.  In every case, assignments are to be typed neatly using proper English in Markdown.

In the last few weeks, we have discussed data structures, plotting using base R graphics, ggplot2, creation of multipanel plots in both plotting paradigms, and use of the formula interface.  

# Problem 1 -- Rstudio.cloud Primers

If you need a bit more work with ggplot2, please see Rstudio.cloud's primers "Visualize Data".

# Problem 2

Now that we have the R environment setup and have a basic understanding of R, let's add Markdown (choose File, New File, R Markdown, pdf).  

Let's go ahead and save the file as is.  Save the file to the directory containing the *README.md* file you created and committed to your git repo in Homework 3.  The filename should be: HW1_pid, i.e. for me it would be HW3_rsettlag.  Note, in future homework assignments I -may-, ok, *will likely* forget to increment the number, please infer the proper assignment number from the assignment...

You will use this new R Markdown file for the remainder of this homework.


# Problem 3

Here we will create a linear model and associated basic diagnostic plots for some COVID data as in the previous homework  As a reminder, what we want is US data only and we will look at the 14 day moving average per 100k people.  We are going to limit the time period to June-July 2020.

```{r libraryLoad, echo=TRUE, include=TRUE, eval=FALSE}
install.packages('data.table')
library(data.table)
covid_raw <- fread("https://opendata.ecdc.europa.eu/covid19/casedistribution/csv")
us <- covid_raw[covid_raw$countriesAndTerritories == 'United_States_of_America',]
us_filtered <- us[us$month %in% c(6:7),]
us_filtered$index <- rev(1:dim(us_filtered)[1]) 
fit<-lm(`Cumulative_number_for_14_days_of_COVID-19_cases_per_100000`~index, data=us_filtered)

## augment the data as previous
install.packages("broom")
fit.diags <- broom::augment(fit)
```

As before, we are going to restrict this to four plots:  

1. residuals vs fitted  
2. normal Q-Q  
3. scale-location  
4. residuals vs leverage  

Create them using ggplot2 functions.

# Problem 4

As before, we want to take the plots of the previous problem and create a single multipanel plot, this time, using ggarrange (ggpubr) or a similar canvas controlling package (gridExtra, patchwork).

**This document containing solutions to Problems 2-4 should be typed in RMarkdown, using proper English, and knitted to create a pdf document.  Do NOT print, we will use git to submit this assignment as detailed below.**

# Problem 5
Please knit this document to PDF (name should be `HW3_pid`) and push to GitHub:

In the R Terminal, type:  

* git pull  
* git add HW3\_pid.[pR]*  (NOTE: this should add two files)  
* git commit -m "final HW3 submission"   
* git push

A more detailed description is on the course website under *Submitting Homework*.

## Reminder on where to find Git help:  
Read through the Git help Chapters 1 and 2.  
<https://git-scm.com/book/en/v2>






