---
title:  'Week 8 - Apply family of functions'
author: "Robert Settlage"
date: "2021-10-19"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
header-includes: \setlength\parindent{24pt} \usepackage{MnSymbol} \usepackage{mathrsfs}
---


```{r setup, eval=T, echo=F, include=F}
  library(ggplot2); library(microbenchmark)
```

## Today's Agenda

- Function optimizing  
- Apply functions
    
## Functions and Timing

As you start to make functions, it is a good idea to think about how long things take.

Really good blog on this here:  <https://www.r-bloggers.com/5-ways-to-measure-running-time-of-r-code/>

Consider the following code and timings:

```{r echo=T, eval=T, cache=F, tidy=F}

    library("microbenchmark")
    ## calculate mean of vector
    lvec <- 10000
    vec <- rnorm(lvec)
    times<-microbenchmark(result1<-mean(vec),
                          result2<-sum(vec)/length(vec),
                          result3<-sum(vec)/lvec,times = 100, unit = "ms")
    identical(result1, result2)
```

## Functions and Timing

```{r echo=T, eval=T, cache=F, tidy=T, include=T, fig.height=4}
    ggplot2::autoplot(times)
```

## Matrix

Last week, we went through matrix operations and hopefully talked about why we don't invert matrices:

<https://www.johndcook.com/blog/2010/01/19/dont-invert-that-matrix/>
<https://www.r-bloggers.com/dont-invert-that-matrix-why-and-how/>

```{r echo=T, eval=T, include=T}
    
    n <- 2000
    A = matrix(rnorm(n*n, mean=0, sd=5), n, n)
    y <- rnorm(n)
    time1 <- system.time({b_inverting <- solve(A)%*%y})
    time2 <- system.time({b_solving <- solve(A,y)})
    time3 <- system.time({b_matrix_solve <- Matrix::solve(A,y)})

    print(rbind(time1, time2, time3)[,1:3])
    max(c(abs(b_inverting - b_solving),abs(b_matrix_solve - b_solving)))
```


## Apply family of functions

We often want to "apply" a function along a "margin" of our data.  In the previous example, we used a matrix operation, but what if we have a custom function we wish to use.

In R, we have helper functions to further simplify our code by obviating the for loop.

Apply family:

apply, lapply , sapply, vapply, mapply, rapply, and tapply

Nice tutorial:  
<https://www.r-bloggers.com/r-tutorial-on-the-apply-family-of-functions/>

## Apply detail

*apply(X, MARGIN, FUN, ...)*

```{r echo=T, eval=F, include=T}
    # ?apply
    x <- cbind(x1 = 3, x2 = c(4:1, 2:5))
    dimnames(x)[[1]] <- letters[1:8]
    x
    apply(x, 2, mean, trim = .2)
    col.sums <- apply(x, 2, sum)
    row.sums <- apply(x, 1, sum)

```

## What is difference between various apply functions

We could start with the help `?apply, ?sapply`. The main differences are:  
 <http://www.dummies.com/programming/r/how-to-use-the-apply-family-of-functions-in-r/>
 

